# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, janitor, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
#Cleaning names
data_ncd <- data_ncd %>% 
  clean_names() 
```

```{r}
table_ncd <- data_ncd %>% 
  rename(dalys=dal_ys_disability_adjusted_life_years_non_communicable_diseases_sex_both_age_age_standardized_rate,country=entity) %>%
  filter(code=="BFA"|code=="ZMB"|code=="CHE") 
# You should pivot the data to show each country in a separate column.
table_ncd_wide <- table_ncd %>%
  pivot_wider(values_from =dalys, names_from =country) 
  # Use kable() from the knitr package to render the table.
 kable(table_ncd_wide) 
 
```

## Plot Showing Trends in NCD Burden Over Time

```{r}
ggplot(data =table_ncd, aes(x = year, y = dalys, color = country)) +
  geom_line() +
  labs(title = "Trend of non communicablel disease burden 1990-2019",
       x = "Year",
       y = "Burden of disease (in dalys",
       color = "country") +
  scale_color_manual(values = c("Burkina Faso" = "blue", "Switzerland" = "red", "Zambia" = "green")) 
```

## Summary of NCD Burden Findings
For Switzerland, there has been a sharp downward trend over the 3 decades. Moreover, the levels of disease burden as much lower than those of Zambia and Burkina Faso, two low-income countries. The Burden has consistently been higher than that of Burkina Faso, up until around 2015 when the two countries converge to the same level of burden (around 25,000 dalys).
