# Overall Disease Burden

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, janitor, here)

# Import the overall disease data
data_overall <- read_csv(here("data", "burden-of-disease-all-causes.csv"))
```

## Table of Estimates for Overall Disease Burden Over Time

```{r}
#Cleaning names
data_overall <- data_overall %>% 
  clean_names() 
```

```{r}
table_overall <- data_overall %>% 
  rename(dalys=dal_ys_disability_adjusted_life_years_all_causes_sex_both_age_age_standardized_rate,country=entity) %>%
  filter(code=="BFA"|code=="ZMB"|code=="CHE") 
# You should pivot the data to show each country in a separate column.
table_overall_wide <- table_overall %>%
  pivot_wider(values_from =dalys, names_from =country) 
  # Use kable() from the knitr package to render the table.
 kable(table_overall_wide)
```

## Plot Showing Trends in Overall Disease Burden Over Time

```{r}
ggplot(data =table_overall, aes(x = year, y = dalys, color = country)) +
  geom_line() +
  labs(title = "Trend of disease burden (all causes) 1990-2019",
       x = "Year",
       y = "Burden of disease (in dalys",
       color = "country") +
  scale_color_manual(values = c("Burkina Faso" = "blue", "Switzerland" = "red", "Zambia" = "green"))
```

## Summary of Overall Disease Burden Findings

#The burden of disease (BoD) due to all causes for Switzerland is extremely low compared to Burkina Faso and Zambia, and has been declining quite consistenly over time. The BoD for Zambia has been very high during two decades (1990s and 2000s), certainly as a result of the HIV epidemic. Around 2012, Zambia BoD decreased over time and became comparable to that of Burkina Faso
